---
title: "Composite Plate PMI"
author: "Zachary del Rosario"
date: 2020-07-14
output:
  github_document
---

This document executes various PMI scenario studies of the angle ply composite
plate. This code was used to generate Figure 4 in the manuscript.

# Setup
<!-- -------------------------------------------------- -->

Setup helper functions to study scenarios below.

```{r setup}
library(tidyverse)
library(rzdr)

file_weights <-  "../data/plate_weights.csv"
file_param <-  "../data/plate_parameters.csv"

repair_tex <- function(string) {
  string %>%
  str_replace_all("DOLLAR", "$") %>%
  str_replace_all("POW", "^") %>%
  str_replace_all("\\\\_", "_") %>%
  str_replace_all("\\\\\\{", "\\{") %>%
  str_replace_all("\\\\\\}", "\\}") %>%
  str_replace_all("rho", "\\\\rho") %>%
  str_replace_all("sigma", "\\\\sigma") %>%
  str_replace_all("nu", "\\\\nu") %>%
  str_replace_all("theta", "\\\\theta")
}
```

```{r load}
df_weights <- read_csv(file_weights)
df_param <- read_csv(file_param)
```

## Document
<!-- ------------------------- -->

Write the parameter table for inclusion in writeup:

```{r write-table-param}
df_units <-
  tribble(
       ~factor,   ~units, ~tex,
        "E1_0",     "Pa", "E_{1,0}",
        "E1_1",     "Pa", "E_{1,1}",
        "E2_0",     "Pa", "E_{2,0}",
        "E2_1",     "Pa", "E_{2,1}",
       "G12_0",     "Pa", "G_{12,0}",
       "G12_1",     "Pa", "G_{12,1}",
      "nu12_0",      "-", "nu_{12,0}",
      "nu12_1",      "-", "nu_{12,1}",
          "Nx",      "N", "N_x",
     "sigma_c",     "Pa", "sigma_c",
         "t_0",      "m", "t_0",
         "t_1",      "m", "t_1",
     "theta_0",      "-", "theta_0",
     "theta_1",      "-", "theta_1",
         "rho", "kg/mPOW3", "rho",
  )

df_param %>%
  left_join(
    df_units,
    by = "factor"
  ) %>%
  mutate(
    mean = exp(mu),
    factor = str_c("DOLLAR", tex, " (", units, ")DOLLAR")
  ) %>%
  select(
    `Property (Units)` = factor,
    Mean = mean,
    COV = sigma
  ) %>%
  knitr::kable(digits = 3, format = "latex") %>%
  repair_tex() %>%
  cat(., file = "../tables/tab-laminate-param.tex")
```

Write the weights table for inclusion in writeup:

```{r write-table-weights}
df_units <-
  tribble(
       ~factor,   ~units, ~tex,
        "E1_0",     "Pa", "E_{1,0}",
        "E1_1",     "Pa", "E_{1,1}",
        "E2_0",     "Pa", "E_{2,0}",
        "E2_1",     "Pa", "E_{2,1}",
       "G12_0",     "Pa", "G_{12,0}",
       "G12_1",     "Pa", "G_{12,1}",
      "nu12_0",      "-", "nu_{12,0}",
      "nu12_1",      "-", "nu_{12,1}",
          "Nx",      "N", "N_x",
     "sigma_c",     "Pa", "sigma_c",
         "t_0",      "m", "t_0",
         "t_1",      "m", "t_1",
     "theta_0",      "-", "theta_0",
     "theta_1",      "-", "theta_1",
         "rho", "kg/mPOW3", "rho",
           "p",      "-",  "p",
           "q",      "-",  "q"
  )

df_weights %>%
  mutate(
    vector = str_match(factor, "^[vw]"),
    factor = str_remove_all(factor, "^[vw]-")
  ) %>%
  left_join(
    df_units,
    by = "factor"
  ) %>%
  mutate(factor = str_c("DOLLAR", tex, "DOLLAR")) %>%
  select(
    Variable = factor,
    Vector   = vector,
    Exponent = value
  ) %>%
  knitr::kable(digits = 3, format = "latex") %>%
  repair_tex() %>%
  cat(., file = "../tables/tab-laminate-weights.tex")
```

## Helper functions
<!-- ------------------------- -->

Unpack the composite plate (angle ply) model weights.

```{r arrange-weights}
## Unpack vectors
v <-
  df_weights %>%
  filter(str_detect(factor, "^v")) %>%
  mutate(factor = str_remove(factor, "v-")) %>%
  pivot_wider(
    names_from = factor,
    values_from = value
  ) %>%
  unlist(.)

w <-
  df_weights %>%
  filter(str_detect(factor, "^w")) %>%
  mutate(factor = str_remove(factor, "w-")) %>%
  pivot_wider(
    names_from = factor,
    values_from = value
  ) %>%
  unlist(.)

## Unpack scalars
p <-
  df_weights %>%
  filter(factor == "p") %>%
  pull(value) %>%
  .[[1]]

q <-
  df_weights %>%
  filter(factor == "q") %>%
  pull(value) %>%
  .[[1]]

## Print values
v
w
p
q
```

The following computes the model (combined) parameters under a specified
distribution parameter scenario `df`.

```{r calc-parameters}
comp_param <- function(df) {
  ## Extract the parameters
  vec_mu <-
    df %>%
    select(factor, mu) %>%
    pivot_wider(
      names_from = factor,
      values_from = mu
    ) %>%
    select(all_of(names(w))) %>%
    unlist()

  vec_sigma <-
    df %>%
    select(factor, sigma) %>%
    pivot_wider(
      names_from = factor,
      values_from = sigma
    ) %>%
    select(all_of(names(w))) %>%
    unlist()

  ## Compute combined parameters
  mu_v <- v %*% vec_mu
  mu_w <- w %*% vec_mu

  sig_v <- sqrt(v^2 %*% vec_sigma^2)
  sig_w <- sqrt(w^2 %*% vec_sigma^2)

  ## Return packaged values
  tibble(mu_v = mu_v, mu_w = mu_w, sig_v = sig_v, sig_w = sig_w)
}

## Test function
comp_param(df_param)
```

The following compute the PMI factors under a specified parameter scenario
`df`, desired reliability `beta_t`, and sampling count `n`.

```{r calc-factors}
comp_mmi <- function(df = df_param) {
  df_param <- comp_param(df)

  mu_w <- df_param %>%
    pull(mu_w) %>%
    .[[1]]
  mu_v <- df_param %>%
    pull(mu_v) %>%
    .[[1]]

  MMI <- exp(p/q * mu_w - mu_v)
  MMI
}

comp_f_beta <- function(beta_t, df = df_param) {
  df_param <- comp_param(df)

  sig_w <- df_param %>%
    pull(sig_w) %>%
    .[[1]]

  f_beta <- exp(p/q * beta_t * sig_w)
}

comp_f_b <- function(beta_t, n, df = df_param, C = 0.95) {
  df_param <- comp_param(df)

  b <- qt(p = C, ncp = sqrt(n) * beta_t, df = n) / sqrt(n) - beta_t

  sig_w <- df_param %>%
    pull(sig_w) %>%
    .[[1]]

  f_beta <- exp(p/q * b * sig_w)
}
```

The following function modifies the distribution parameters by reducing the
variability of material properties only; this models an improvement in
manufacturing process control.

```{r reduce-helper}
reduce_sig <- function(reduce) {
  ## Reduce the variability (sigma) of the material properties by a given factor
  df_param %>%
    mutate(
      sigma = case_when(
        factor %in% c("Nx", "t_0", "t_1", "theta_0", "theta_1") ~ sigma,
        TRUE ~ sigma * (1 - reduce)
      )
    )
}
```

# Scenarios
<!-- -------------------------------------------------- -->

## Sample Sizes
<!-- ------------------------- -->

Let's consider a range of sample sizes.

```{r sample-sweep}
df_res <-
  map_dfr(
    10^seq(1, 3, length.out = 10),
    function(n) {
      mmi <- comp_mmi()
      f_beta <- comp_f_beta(beta_t = 3)
      f_b <- comp_f_b(beta_t = 3, n = n)

      tibble(mmi = mmi, f_beta = f_beta, f_b = f_b, n = n)
    }
  )
df_res
```

Visualize the scenarios.

```{r vis-sweep}
df_res %>%
  mutate(
    knockdown_Beta = 1 - f_beta,
    ## knockdown_b = 1 - f_b,
    knockdown_Total = 1 - f_beta * f_b
  ) %>%
  pivot_longer(
    names_to = "factor",
    names_prefix = "knockdown_",
    values_to = "knockdown",
    cols = matches("knockdown_")
  ) %>%

  ggplot(aes(n, knockdown, color = factor)) +
  geom_line() +
  scale_x_log10() +
  scale_y_continuous(labels = scales::percent) +
  scale_color_discrete(name = "Factor") +
  theme_common() +
  labs(
    x = "Samples",
    y = "Knockdown"
  )

ggsave("../images/plate_sampling_scenarios.eps")
```

**Observations**:

- At the minimum ($n = 10$) the total knockdown is $\approx50\%$
- Sampling at $n = 100$ reduces the total knockdown to $\approx40\%$
- Sampling much beyond $n = 100$ has diminishing returns, as the inherent knockdown (beta factor) dominates

## Sampling vs Process Control
<!-- ------------------------- -->

```{r sample-process-sweep}
scenario_reduced <- function(reduce) {
  df_red <- reduce_sig(reduce)

  map_dfr(
    10^seq(1, 3, length.out = 10),
    function(n) {
      mmi <- comp_mmi(df = df_red)
      f_beta <- comp_f_beta(df = df_red, beta_t = 3)
      f_b <- comp_f_b(df = df_red, beta_t = 3, n = n)

      tibble(mmi = mmi, f_beta = f_beta, f_b = f_b, n = n)
    }
  ) %>%
  mutate(reduce = reduce)
}

df_sweep_reduce <-
  map_dfr(
    c(0, 0.10, 0.25, 0.5),
    scenario_reduced
  )
```

Visualize the scenarios

```{r vis-reduce}
df_sweep_reduce %>%
  mutate(
    knockdown_Beta = 1 - f_beta,
    knockdown_Total = 1 - f_beta * f_b
  ) %>%
  pivot_longer(
    names_to = "factor",
    names_prefix = "knockdown_",
    values_to = "knockdown",
    cols = matches("knockdown_")
  ) %>%
  mutate(
    factor = fct_relevel(factor, "Total", "Beta"),
    r_percent = str_c(as.integer(reduce * 100), "%"),
    r_percent = fct_reorder(r_percent, reduce)
  ) %>%

  ggplot(aes(n, knockdown)) +
  geom_line(
    aes(linetype = factor, color = as.factor(r_percent)),
    size = 2
  ) +
  geom_point(
    data = . %>%
      filter(
        as.integer(n) == 27,
        reduce == 0.5,
        factor == "Total"
      ),
    size = 4
  ) +
  scale_color_discrete(name = "Variability Reduction") +
  scale_linetype(name = "Factor") +

  scale_x_log10() +
  scale_y_continuous(labels = scales::percent) +
  theme_common() +
  theme(legend.position = "bottom", legend.box = "vertical") +
  labs(
    title = "Laminate Angle-Ply Log-Precision Materials Indices",
    x = "Number of Physical Tests",
    y = "Knockdown"
  )

ggsave("../images/plate_reduce_scenarios.eps")
```

Extract the choosen scenario

```{r point-scenario}
df_sweep_reduce %>%
  filter(
    as.integer(n) == 27,
    reduce == 0.5
  ) %>%
  mutate(Total = 1 - f_beta * f_b) %>%
  select(Total, n, reduce)

```
