---
title: "Laminate Plate Modeling"
author: "Zach del Rosario"
date: 2020-07-13
output:
  pdf_document
---

**Objective**: If feasible, fit a monomial-standard model of the angle ply
problem.

**Strategy**: Log-transforming the data will turn products into linear
combinations. If we can successfully log-transform all the inputs, we can
simply perform linear regression to fit the exponents.

Technically there are two decision variables in this problem: the individual
thicknesses of the two-layer composite. We'll fit on these two factors
individually, but fix a single thickness in the final model definition.

# Setup
<!-- -------------------------------------------------- -->

```{r setup}
library(tidyverse)

file_sim <- "../data/plate_sim.csv"
file_weights <-  "../data/plate_weights.csv"
file_param <-  "../data/plate_parameters.csv"
```

Note that the plane laminate implements its "g" as $g = 1 - \sigma_a /
\sigma_c$. Let's upfront back-calculate the applied stress.

```{r load}
df_sim <-
  read_csv(file_sim) %>%
  mutate(
    app_22_t_0 = sigma_22_t_0 * (1 - g_22_tension_0),
    app_22_t_1 = sigma_22_t_1 * (1 - g_22_tension_1)
  )
```

I only plan to use the tensile limit states, so I leave the other applied
stresses uncalculated.

# EDA
<!-- -------------------------------------------------- -->

## Zero-crossings
<!-- ------------------------- -->

Are there any variables that cross zero?

```{r zero-crossing}
df_sim %>%
  summarize(across(everything(), ~ any(. < 0) & any(0 < .))) %>%
  pivot_longer(
    names_to = "factor",
    values_to = "crossing",
    cols = everything()
  ) %>%
  filter(crossing)
```

**Observations**:

- The fiber-orthogonal failure mode is active, as accounted by the limit state crossings. We will deal with these zero-crossing by filtering out negative values.
- The RV model to allows negative poisson values. That's not technically wrong (according to thermodynamic admissibility), but still strange.
- Some applied stresses along the 22 direction are negative (compressive). These are highly non-critical, but strange.

How many poisson values go negative?

```{r negative-poisson}
df_sim %>%
  select(nu12_0, nu12_1) %>%
  pivot_longer(
    names_to = "factor",
    values_to = "value",
    cols = everything()
  ) %>%

  ggplot(aes(value)) +
  geom_histogram() +
  facet_grid(~factor)
```

There's a very small tail crossing zero. Let's truncate this.

Let's inspect the negative-stress cases. What can explain those?

```{r negative-applied-stress}
df_sim %>%
  ggplot(aes(theta_0, theta_1)) +
  geom_point() +
  geom_point(
    data = . %>%
      filter(app_22_t_0 < 0),
    color = "red"
  )
```

Interesting; at strong off-nominal ply angles the cross-fiber stress can become compressive. This makes geometric sense.

There's not much danger in simply truncating these cases; this instances are far from the failure state.

## Negative factors
<!-- ------------------------- -->

Are there any factors that are negative?

```{r negative-factors}
df_sim %>%
  summarize(across(everything(), ~ all(. < 0))) %>%
  pivot_longer(
    names_to = "factor",
    values_to = "crossing",
    cols = everything()
  ) %>%
  filter(crossing)
```

The only negative factor is `theta_0`. We can re-code this variable as positive.

## Data processing
<!-- ------------------------- -->

Let's implement the choices above:

- Filter the negative poisson values
- Re-code `theta_0` as negative
- Log-transform the factors and response
- Compute the applied stress $\sigma_a = \sigma_c - g$

```{r process}
var_factors <-
  df_sim %>%
  select(-contains("g_")) %>%
  select(
    -contains("sigma_11"),
    -contains("sigma_12"),
    -contains("22_c"),
    -contains("app")
  ) %>%
  names()

df_wk <-
  df_sim %>%
  filter(
    nu12_0 > 0,
    nu12_1 > 0,
    app_22_t_0 > 0,
    app_22_t_1 > 0
  ) %>%
  mutate(theta_0 = -theta_0) %>%
  select(-contains("g_")) %>%
  mutate(across(everything(), log))
```

# Modeling
<!-- -------------------------------------------------- -->

## Function Model
<!-- ------------------------- -->

Successful operation corresponds to

$$\sigma_a / \sigma_c < 1$$.

Model the applied stress $\sigma_a$; we already know the exponent for
$\sigma_c$ (negative one).

Let's fit two models: One for each of the two cross-ply failure modes. Filter
out the irrelevant strength values. Otherwise fit on all factors to account for
cross-ply stress interactions.

```{r fit-applied_22_0}
df_tmp0 <-
  df_wk %>%
  select(all_of(var_factors), app_22_t_0) %>%
  select(-contains("sigma"))

fit_22_0 <-
  df_tmp0 %>%
  lm(
    formula = app_22_t_0 ~ .,
    data = .
  )

summary(fit_22_0)
```

The fit for `app_22_t_0` is decent.

```{r fit-applied_22_1}
df_tmp1 <-
  df_wk %>%
  select(all_of(var_factors), app_22_t_1) %>%
  select(-contains("sigma"))

fit_22_1 <-
  df_tmp1 %>%
  lm(
    formula = app_22_t_1 ~ .,
    data = .
  )

summary(fit_22_1)
```

The fit for `app_22_t_1` is decent.

Compare the two model coefficients; are they symmetric?

```{r fit-compare}
df_weights <-
  bind_rows(
    fit_22_0 %>%
    broom::tidy() %>%
    mutate(model = 0),
    fit_22_1 %>%
    broom::tidy() %>%
    mutate(model = 1)
  ) %>%
  select(term, estimate, model) %>%
  pivot_wider(
    names_from = model,
    names_prefix = "estimate_",
    values_from = estimate,
    values_fill = 0
  ) %>%
  arrange(term)

df_weights
```

**Observations**:

- The weights are roughly symmetric across the plies. We would be justified in using either of the two models.

## Monomial-standard form
<!-- ------------------------- -->

```{r make-weights}
## Reliability constraint
w <-
  bind_rows(
    df_weights %>%
    select(term, weight = estimate_0) %>%
      filter(
        term != "(Intercept)",
        !str_detect(term, "t_"),
        !str_detect(term, "theta_"),
        term != "Nx"
      ),
    tribble(
      ~term, ~weight,
      "sigma_c", -1,
      "rho", 0
    )
  ) %>%
  pivot_wider(
    names_from = term,
    values_from = weight
  ) %>%
  unlist(.)

q <-
  df_weights %>%
  filter(str_detect(term, "t_")) %>%
  summarize(t = sum(estimate_0)) %>%
  pull(t) %>%
  .[[1]]

## Objective
v <-
  tibble(term = names(w)) %>%
  mutate(weight = if_else(term == "rho", 1, 0)) %>%
  pivot_wider(
    names_from = term,
    values_from = weight
  ) %>%
  unlist(.)

p <- 1
```

Show the weights

```{r show-weights}
## Objective
p
v

## Reliability constraint
q
w
```

Save the weights to a file

```{r save-weights}
df_all_weights <-
  bind_rows(
    tibble(
      value = v,
      factor = names(v)
    ) %>%
    mutate(factor = str_c("v-", factor)),
    tibble(
      value = w,
      factor = names(w)
    ) %>%
    mutate(factor = str_c("w-", factor))
  ) %>%
  bind_rows(
    tibble(value = p, factor = "p")
  ) %>%
  bind_rows(
    tibble(value = q, factor = "q")
  )

write_csv(df_all_weights, file_weights)

df_all_weights %>%
  knitr::kable()
```

## Density Model
<!-- ------------------------- -->

A joint density is already defined in the plate model; use the simulations to
fit parameter values for a lognormal distribution. Use the working
log-transformed data.

```{r set-parameters}
df_parameters <-
  ## Collapse the per-ply data
  df_wk %>%
  select(all_of(var_factors)) %>%
  pivot_longer(
    names_to = "factor",
    values_to = "value",
    cols = matches("sigma")
  ) %>%
  mutate(
    factor = str_remove(factor, "_\\d$"),
    factor = str_replace(
      factor,
      "sigma_22_t",
      "sigma_c"
    )
  ) %>%
  pivot_wider(
    names_from = factor,
    values_from = value
  ) %>%
  unnest(-Nx) %>%
  ## Estimate parameters
  summarize(across(
    everything(),
    c("mu" = mean, "sigma" = ~ sd(.))
  )) %>%
  pivot_longer(
    names_to = c("factor", ".value"),
    names_pattern = "(.*)_(mu|sigma)",
    cols = everything()
  ) %>%
  arrange(factor) %>%
  bind_rows(tibble(factor = "rho", mu = log(0.05), sigma = 0.01))

write_csv(df_parameters, file_param)

df_parameters
```

We will use these as ground-truth parameters for the subsequent PMI study.
